```{r setoptions, echo=FALSE}
library(knitr)
library(ggplot2)
opts_chunk$set(echo = TRUE)
```

## Analysis of the ToothGrowth data in the R datasets package. 

1. Load the ToothGrowth data and perform some basic exploratory data analyses 

### Plotting the data
```{r, fig.width=6, fig.height=3, fig.align='center'}
library(ggplot2)
id<-factor(c(rep(1:10,3),rep(11:20,3)))
tooth<-cbind(ToothGrowth, id=id)
g <- ggplot(tooth, aes(x = dose, y = len, group = factor(id)))
g <- g + facet_grid(. ~ supp)
g <- g + geom_line( size = 1, aes(colour = id)) + geom_point(size =10, pch = 21, fill = "salmon", alpha = .5)
g
```

2. Provide a basic summary of the data.

```{r}
str(ToothGrowth)
```

```{r}
head(ToothGrowth)
```

```{r}
summary(ToothGrowth)
```

3. Use confidence intervals and hypothesis tests to compare tooth growth by supp and dose. (Use the techniques from class even if there's other approaches worth considering)
```{r}
vcg1<-tooth$len[1:10]
vcg2<-tooth$len[11:20]
vcg3<-tooth$len[21:30]

difference<-vcg3-vcg1
mn<-mean(difference)
s<-sd(difference)
n<-10

mn + c(-1,1)*qt(.975, n-1)*s/sqrt(n)
t.test(difference)
```

For Supp VC we reject the null hypotesis with p `r t.test(difference)$p.value`

```{r}
vcg1<-tooth$len[31:40]
vcg2<-tooth$len[41:50]
vcg3<-tooth$len[51:60]

difference<-vcg3-vcg1
mn<-mean(difference)
s<-sd(difference)
n<-10

mn + c(-1,1)*qt(.975, n-1)*s/sqrt(n)
t.test(difference)
```

For Supp OJ we reject the null hypotesis with p `r t.test(difference)$p.value`

4. State your conclusions and the assumptions needed for your conclusions. 

Conclusion: Teeth grew with both Supp.  However, teeth grew more on average with Supp VC than with Supp OJ.  
Assumptions: It was assumed that there were two groups of 10 guinea pigs (one group on each Supp) and that the data within in Supp group was paired.